Proposal outline/notes for DPG and Compressible Navier-Stokes

### 2012-06-15 docs/Jesse/Proposal/DPGrobustness.tex
adding more additions (NL DPG) to proposal

\section{Model problem and robustness}
\seclab{sec:modelSec}
The remainder of the paper focuses on a convection-diffusion model problem using the abstract theory that we
have discussed so far. In particular, we shall use the DPG method based on the ultra-weak formulation with optimal test functions to solve the model problem and analyze its behavior with respect to $\epsilon$. 
Our goal is to show the robustness of the method with respect to $\epsilon$, and demonstrate its usefulness as  a numerical approach to solving singular-perturbed problems.

We consider the following model convection-diffusion problem on a domain
$\Omega \subset \mathbb{R}^d$ with boundary $\pO \equiv \Gamma$
\begin{equation}
\div (\beta u) - \epsilon\Delta u = f  \in \L \label{primal},
\end{equation}
which can be cast into the first order form on the group variable
$\LRp{u,\sigma}$ as
\begin{equation}
\eqnlab{CDstrong}
A \LRp{u,\sigma} \coloneqq \LRs {
\begin{array}{c}
\div (\beta u - \sigma) \\ \frac{1}{\epsilon}\sigma - \grad u
\end{array}} = \LRs{
\begin{array}{c}
f \\ 0
\end{array}
}.
\end{equation}
Using the abstract ultra-weak formulation developed in Section
\secref{abstractUweak} for the first order system of PDEs \eqnref{CDstrong} we
obtain
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\div \tau - \beta \cdot \grad
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad v\right)_{\Oh} - \LRa{
\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{ \widehat{f}_n,
  \jump{v} }_{\Gh},
\end{align*}
where $\LRp{v, \tau}$ is the group test function. It should be pointed
out that the divergence and gradient operators are understood to act
element-wise on test functions $\LRp{v, \tau}$ in the broken graph
space $ D\LRp{A_h^*} \coloneqq  H^1(\Oh) \times H({\rm div}, \Oh)$, but
globally as usual on conforming test functions, i.e.\ $ \LRp{v, \tau}
\in  H^1(\Omega) \times H({\rm div}, \Omega)$. It follows that the
canonical test norm can be written as
\[
\|\left(v, \tau\right)\|_V^2 = \|\left(v, \tau\right)\|_{H^1(\Oh) \times H({\rm div},\Oh)}^2
= \sum_{K\in \Oh} \|\left(v, \tau\right)\|_{H^1(K) \times H({\rm
    div},K)}^2,
\]
where
\[
\|\left(v, \tau \right)\|_{H^1(K) \times H({\rm div},K)}^2 =
\|v\|_{L^2(K)}^2 + \|\grad v\|_{L^2(K)}^2 + \|\tau\|_{L^2(K)}^2 +
\|\div \tau\|_{L^2(K)}^2.
\]

In order to define the proper norm on the trial space, boundary
conditions need to be specified. We begin by splitting the boundary
$\Gamma$ as follows
\begin{align*}
\Gamma_{-} &\coloneqq \{x\in \Gamma; \beta_n(x) < 0\}, \quad {\rm
  (inflow)}\\ \Gamma_{+} &\coloneqq \{x\in \Gamma; \beta_n(x) > 0\},
\quad {\rm (outflow)}\\ \Gamma_{0} &\coloneqq \{x\in \Gamma;
\beta_n(x) = 0\},
\end{align*}
where $\beta_n \coloneqq \beta \cdot n$.
Previous work in \cite{DPGrobustness} adopted Dirichlet boundary
conditions everywhere on $\Gamma$.  
%and we could obtain the robustness only with special weighted norms in which the weight varied spatially. 
In this paper, we employ the inflow
condition of Hesthaven \etal \cite{Hesthaven96astable}, where we set
\[
\beta_n u - \sigma_n  = u_0, \quad \text{on } \Gm,
\]
instead of $\beta_n u = u_0$. The former resembles the latter
as $\epsilon$ approaches zero\footnote{For our model problem, as for many problems of interest in computational fluid dynamics, we expect $\grad u$ to be small near the inflow, and that the solutions to \eqref{primal} using $\beta_n u - \sigma_n = f_n = u_0$ on $\Gamma_-$ will converge to that using $u=u_0$ on $\Gamma_-$  for sufficiently small $\epsilon$.}; however, the latter induces a more ``well-behaved" adjoint problem than the former, which, as we will discuss, affects the performance of DPG.

On the outflow boundary, we apply standard homogeneous Dirichlet boundary conditions
\[
u = 0, \quad \text{on } \Gp.
\]
This paper is intended to act as an extension of work presented by Heuer and Demkowicz in \cite{DPGrobustness}. The primary focus of the paper is to analyze the DPG method and extend previous results under this new choice of inflow boundary conditions. The difference in the performance of DPG under both new and old boundary conditions is connected to the difference in the adjoint problems induced under each boundary condition. The secondary contribution of this paper will be to analyze the performance of DPG under a new mesh-dependent test norm.

@@

\section{Analysis of a DPG test norm}
\seclab{sec:testNormSec}
We are interested in computing DPG optimal test functions for the convection-diffusion equation with very small values of $\epsilon$; due to the difficulty of approximating optimal test functions, we conclude that the use of the quasi-optimal test norm is infeasible towards this goal.

@@

\subsection{Decomposition into analyzable components}
\label{sec:strategy2}

@@

\subsection{Adjoint estimates}
\label{sec:strategy3}

@@

%Together, these estimates will be sufficient to construct a complete norm on $V$ and demonstrate bounds on induced test norms for different scalings of norms on $U$.  
%\tanbui{}{again, using the same notation
%  $(v, \tau)$ for all three problems is not a good idea. Why don't we
%  use $(v_i,\tau_i)$ introduced above? Things are much clearer this
%  way?}
\subsection{A mesh-dependent test norm}
\seclab{sec:meshDep}
Ideally, we would be interested in the use of the test norm 
\[
\|\left(v,\tau\right)\|_{V}^2 = \|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau\|^2 + \frac{1}{\epsilon}\|\tau\|^2
\]
for practical computations. However, the presence of the term $\|v\|$ together with $\sqrt{\epsilon}\|\grad v\|$ (and similarly $\|\div \tau\|$ and $\frac{1}{\sqrt{\epsilon}}\|\tau\|$ terms) induces boundary layers in the optimal test functions for under-resolved meshes. We can see this by recovering the strong form of the variational problem defining test functions. We first note that the variational problems for the $v$ and $\tau$ components of optimal test functions decouple from each other under this test norm. Then, examining the variational problem for the $v$ component only of an optimal test function, and assuming $\div \beta = 0$ for illustrative purposes, we have 
\begin{align*}
\left(\left(v,0\right),\left(\delta v,\delta\tau\right)\right)_{V} &= (v,\delta v) + \epsilon \left(\grad v,\grad \delta v\right) + \left(\beta \cdot \grad v, \beta \cdot \grad \delta v\right)\\
&= \left(v - \epsilon \Delta v - \div \left(\left(\beta \otimes \beta\right) \grad v\right), \delta v\right)_{L^2} + \langle \epsilon \grad v\cdot n, \delta v\rangle + \langle n\cdot \left(\beta\otimes\beta\right) \grad v , \delta v\rangle.
\end{align*}
After integration by parts, we recover the strong form of the operator $L$ inducing such a variational problem 
\[
Lv \coloneqq v - \epsilon \Delta v - \div \left(\left(\beta \otimes \beta\right) \grad v\right),
\]
where we neglect the resulting boundary terms from integration by parts for now.

The streamline direction $\beta$ induces an anisotropic diffusion, while the $\sqrt{\epsilon}\|\grad v\|_{L^2}$ term induces a small isotropic diffusion contribution everywhere. Since any vector in the cross-stream direction is in the null space of the anisotropic diffusion tensor, in the cross-stream directions, the optimal test function is governed only by the cross-stream part of the operator $L$
\[
L_{\rm \beta^\perp} \coloneqq v - \epsilon \Delta v,
\]
and can develop boundary layers in those directions. The presence of boundary layers has been verified through numerical computation as well; using an $H^1$-conforming finite element code with $hp$-adaptivity \cite{demkowicz2006computing}, the solution to the variational problem defining the optimal test function under the above test norm was computed. Figure~\ref{fig:boundaryTest} shows the result of such a computation for the $v$ component of an optimal test function under the above test norm. 
\begin{figure}[!h]
\centering
\subfigure{\includegraphics[scale=.22]{figs/mesh1.png}}
\subfigure{\includegraphics[scale=.22]{figs/sol1.png}}
\caption{The $v$ component of the optimal test function corresponding to flux $\widehat{u} = x(1-x)$ on the bottom side of a unit element for $\epsilon = 0.01$. The corresponding $hp$-mesh used to compute the solution is displayed to the left.}
\label{fig:boundaryTest}
\end{figure}
To avoid boundary layers in the optimal test functions, we follow \cite{DPGrobustness} in scaling the $L^2$ contributions of $v$ by $C_v(K)$, such that, when transformed to the reference element, both $C_v(K)\|v\|^2$ and $\epsilon\|\grad v\|^2$ are of the same magnitude. Similarly, we scale the $L^2$ contributions of $\tau$ by $C_\tau(K)$ such that $\frac{C_\tau(K)}{\epsilon} \|\tau\|^2$ and $\|\div \tau\|^2$ are of the same magnitude as well. For this paper, we consider only isotropic refinements on quadrilateral elements in 2D.

@@

\subsection{Equivalence of energy norm with $\nor{\cdot}_U$}
%\subsection{Relations between trial and test constants}
\label{sec:main_bounds}

The main theoretical result of this paper can now be given:
\begin{lemma}
Under the mesh-dependent test norm
\[
\|\left(v,\tau\right)\|_{V,\Oh}^2 = \|C_vv\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau\|^2 + \|C_\tau\tau\|^2,
\]
where $C_v, C_{\tau}\in L^2(\Omega)$ are defined elementwise through
\begin{align*}
\left.C_v\right |_K &= \min\left\{\sqrt{\frac{\epsilon}{|K|}},1\right\}\\
\left.C_{\tau}\right |_K &= \min\left\{\frac{1}{\sqrt{\epsilon}},\frac{1}{\sqrt{|K|}}\right\}.
\end{align*}
If $\beta$ satisfies \eqref{a_req}, \eqref{b_req}, and \eqref{c_req}, the DPG energy norm $\nor{\cdot}_E$ satisfies the following equivalence relations
\begin{align*}
\nor{u}_{L^2} + \nor{\sigma}_{L^2} + \epsilon\nor{\widehat{u}} + \sqrt{\epsilon}\nor{\widehat{f}_n} 
&\lesssim \nor{\left(u,\sigma, \widehat{u},\widehat{f}_n\right)}_E \\ 
%\nor{\left(u,\sigma, \widehat{u},\widehat{f}_n\right)}_E &\lesssim \nor{u}_{L^2} + \nor{C_\sigma \sigma}_{L^2} + \frac{1}{\sqrt{\epsilon}} \left(\nor{\widehat{u}} + \nor{\widehat{f}_n}\right),
\nor{\left(u,\sigma, \widehat{u},\widehat{f}_n\right)}_E &\lesssim \nor{u}_{L^2} + \nor{\frac{1}{\epsilon C_\tau} \sigma}_{L^2} + \frac{1}{\sqrt{\epsilon}} \left(\nor{\widehat{u}} + \nor{\widehat{f}_n}\right).
\end{align*}
%where $C_\sigma \in L^2(\Omega)$ is defined elementwise through $\left.C_\sigma\right|_K  = \max\left\{ \frac{1}{\sqrt{\epsilon}} , \frac{\sqrt{|K|}}{\epsilon} \right\}$.
\end{lemma}

@@

\subsection{Comparison of boundary conditions}

@@

The adjoint problem corresponding to Lemma~\ref{lemma_stream} in Section~\ref{sec:strategy3} is likewise reduced in 1D to the scalar equation
\begin{align}
\epsilon v_1'' + \beta v_1' &= -g \label{adjoint_1D}
\end{align}
with $v_1\in H^1_0\left((0,1)\right)$.  After multiplying this equation by $\beta v_1'$ and integrate by parts over $\Oh$, we can apply Young's inequality to get
%\[
%\int_0^1 \epsilon v_1'' \beta v_1' + \|\beta v_1'\|_{L^2}^2 = -\int_0^1 g \beta v_1'.\
%]
%Integrating by parts and Young's inequality gives
\begin{align*}
\left.\frac{\epsilon}{2} \beta v_1'^2\right|_0^1 + \|\beta v_1'\|_{L^2}^2 &\leq \frac{1}{2}\|g\|^2 + \frac{1}{2}\|\beta v_1'\|^2,
\end{align*}
implying that
\begin{align*}
\|\beta v_1'\|_{L^2}^2 & \lesssim \|g\|^2 + {\beta \epsilon} v_1'(0)^2.
\end{align*}
Let us restrict ourselves to the cases where $v_1$ is sufficiently smooth for $v'(0)$ to be well defined.  Taking $g=1$ (corresponding to a piecewise constant approximation) we can solve \eqref{adjoint_1D} exactly. 
\begin{figure}[!h]
\centering
\includegraphics[scale=.5]{figs/testBoundary1D.png}
\caption{$v_1(x) = \frac{e^{-\frac{x}{\epsilon}}}{e^{\frac{1}{\epsilon}}-1}\left(e^{\frac{1}{\epsilon}}\left(e^{\frac{x}{\epsilon}}-1\right) + \left(e^{\frac{1}{\epsilon}}-1\right)e^{\frac{x}{\epsilon}}x\right)$, the solution to the adjoint equation for $f=0$ and constant $\beta$ and load $g$ for $\epsilon = .01$.}
\label{fig:testBoundary1D}
\end{figure}
The solution $v_1$ is plotted in Figure~\ref{fig:testBoundary1D}, where we can see that $v_1(x)$ develops strong boundary layers of width $\epsilon$ near the inflow boundary $x=0$. Consequently, $\frac{\epsilon}{2} v_1'(0)^2 \approx \epsilon^{-1}$. Thus, we cannot conclude $\|\beta v'\| \lesssim \|g\|$ when $g$ is a constant,\footnote{Unlike the case of Dirichlet boundary conditions, the inflow condition on $ \widehat{f}_n = u(0)-\epsilon u'(0)$ induces an adjoint boundary condition $\tau(0)=0$, or equivalently $v'(0) = 0$, removing the non-robust term from the estimate.} and as a consequence cannot conclude that the robust error bound $\|u-u_h\|_{L^2} \lesssim \|(u,\sigma,\widehat{u},\widehat{f}_n)-(u_h,\sigma_h,\widehat{u}_h,\widehat{f}_{n,h})\|_E$ holds for the solution $u_h$.  More detailed 1D error bounds for Dirichlet boundary conditions are provided in \cite{DPG3}, and indicate the same lack of robustness under the test norm derived in this paper.\footnote{Demkowicz and Heuer proved in \cite{DPGrobustness} that for Dirichlet boundary conditions, robustness as $\epsilon \rightarrow 0$ is achieved by the test norm
\[
\|\left(\tau, v\right)\|_{V,w}^2 = \|v\| + \epsilon \|\grad v\| + \|\beta \cdot \grad v\|_{w+\epsilon} + \| \div \tau\|_{w+\epsilon} + \frac{1}{\epsilon}\|\tau\|_{w+\epsilon}
\]
where $\|\cdot \|_{w+\epsilon}$ is a weighted $L^2$ norm, where the weight $w \in (0,1)$ is required to vanish on $\Gamma_-$ and satisfy $\grad w = O(1)$. The need for this weight is necessary to account for the loss of robustness at the inflow.}

@@

\begin{figure}[h!]
\centering
\subfigure{
\includegraphics[scale=.38]{figs/roundoff_rates.png}
}
\subfigure{
\includegraphics[scale=.38]{figs/roundoff_ratio.png}
}
\caption{Energy error and $L^2$/energy error ratio for $\epsilon=1e-5$, $\epsilon=1e-6$, $\epsilon=1e-7$. Non-monotonic behavior of the energy error indicates conditioning issues and roundoff effects.}
\label{roundoff_figures}
\end{figure}

@@

\begin{figure}[h!]
\centering
\subfigure{
\includegraphics[scale=.38]{figs/rates_disc.png}
}
\subfigure{
\includegraphics[scale=.38]{figs/ratio_disc.png}
}
\caption{$L^2$ and energy errors, and their ratio for $\epsilon=.01$, $\epsilon=.001$, $\epsilon=.0001$, with discontinuous $u_0$ approximated by a Fourier expansion. }
\label{disc_sol_fourier}
\end{figure}

>>>>

\section{DPG formulation for convection-diffusion}

\seclab{sec:modelSec}
The remainder of this chapter focuses on the convection-diffusion problem using the abstract theory that we
have discussed so far. In particular, we shall use the DPG method based on the ultra-weak formulation with optimal test functions to solve this model problem and analyze its behavior with respect to $\epsilon$. 
Our goal is to show the robustness of the method with respect to $\epsilon$, and demonstrate its usefulness as  a numerical approach to solving singular-perturbed problems.

The convection-diffusion problem is given on a domain
$\Omega \subset \mathbb{R}^d$ with boundary $\pO \equiv \Gamma$
\begin{equation}
\div (\beta u) - \epsilon\Delta u = f  \in \L \label{primal},
\end{equation}
which can be cast into the first order form on the group variable
$\LRp{u,\sigma}$ as
\begin{equation}
\eqnlab{CDstrong}
A \LRp{u,\sigma} \coloneqq \LRs {
\begin{array}{c}
\div (\beta u - \sigma) \\ \frac{1}{\epsilon}\sigma - \grad u
\end{array}} = \LRs{
\begin{array}{c}
f \\ 0
\end{array}
}.
\end{equation}
Using the abstract ultra-weak formulation developed in Section
\secref{abstractUweak} for the first order system of PDEs \eqnref{CDstrong} we
obtain
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\div \tau - \beta \cdot \grad
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad v\right)_{\Oh} - \LRa{
\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{ \widehat{f}_n,
  \jump{v} }_{\Gh},
\end{align*}
where $\LRp{v, \tau}$ is the group test function. It should be pointed
out that the divergence and gradient operators are understood to act
element-wise on test functions $\LRp{v, \tau}$ in the broken graph
space $ D\LRp{A_h^*} \coloneqq  H^1(\Oh) \times H({\rm div}, \Oh)$, but
globally as usual on conforming test functions, i.e.\ $ \LRp{v, \tau}
\in  H^1(\Omega) \times H({\rm div}, \Omega)$. It follows that the
canonical test norm can be written as
\[
\|\left(v, \tau\right)\|_V^2 = \|\left(v, \tau\right)\|_{H^1(\Oh) \times H({\rm div},\Oh)}^2
= \sum_{K\in \Oh} \|\left(v, \tau\right)\|_{H^1(K) \times H({\rm
    div},K)}^2,
\]
where
\[
\|\left(v, \tau \right)\|_{H^1(K) \times H({\rm div},K)}^2 =
\|v\|_{L^2(K)}^2 + \|\grad v\|_{L^2(K)}^2 + \|\tau\|_{L^2(K)}^2 +
\|\div \tau\|_{L^2(K)}^2.
\]

In order to define the proper norm on the trial space, boundary
conditions need to be specified. We begin by splitting the boundary
$\Gamma$ as follows
\begin{align*}
\Gamma_{-} &\coloneqq \{x\in \Gamma; \beta_n(x) < 0\}, \quad {\rm
  (inflow)}\\ \Gamma_{+} &\coloneqq \{x\in \Gamma; \beta_n(x) > 0\},
\quad {\rm (outflow)}\\ \Gamma_{0} &\coloneqq \{x\in \Gamma;
\beta_n(x) = 0\},
\end{align*}
where $\beta_n \coloneqq \beta \cdot n$.
Previous work in \cite{DPGrobustness} adopted Dirichlet boundary
conditions everywhere on $\Gamma$.  
%and we could obtain the robustness only with special weighted norms in which the weight varied spatially. 
We employ instead the inflow
condition of Hesthaven \etal \cite{Hesthaven96astable}, where we set
\[
\beta_n u - \sigma_n  = u_0, \quad \text{on } \Gm,
\]
instead of $\beta_n u = u_0$. The former resembles the latter
as $\epsilon$ approaches zero\footnote{For our model problem, as for many problems of interest in computational fluid dynamics, we expect $\grad u$ to be small near the inflow, and that the solutions to \eqref{primal} using $\beta_n u - \sigma_n = f_n = u_0$ on $\Gamma_-$ will converge to that using $u=u_0$ on $\Gamma_-$  for sufficiently small $\epsilon$.}; however, the latter induces a more ``well-behaved" adjoint problem than the former, which, as we will discuss, affects the performance of DPG.

On the outflow boundary, we apply standard homogeneous Dirichlet boundary conditions
\[
u = 0, \quad \text{on } \Gp.
\]
This work is intended to act as an extension of work presented by Heuer and Demkowicz in \cite{DPGrobustness}. The primary focus of this chapter is to analyze the DPG method and extend previous results under this new choice of inflow boundary conditions. The difference in the performance of DPG under both new and old boundary conditions is connected to the difference in the adjoint problems induced under each boundary condition. The secondary contribution of this work will be to analyze the performance of DPG under a new mesh-dependent test norm.

@@

\subsection{Analysis of a DPG test norm}
\seclab{sec:testNormSec}
We are interested in computing DPG optimal test functions for the convection-diffusion equation with very small values of $\epsilon$; due to the difficulty of approximating optimal test functions, we conclude that the use of the quasi-optimal test norm is infeasible towards this goal.

@@

\subsubsection{Decomposition into analyzable components}
\label{sec:strategy2}

@@

\subsubsection{Adjoint estimates}
\label{sec:strategy3}

@@

%Together, these estimates will be sufficient to construct a complete norm on $V$ and demonstrate bounds on induced test norms for different scalings of norms on $U$.  
%\tanbui{}{again, using the same notation
%  $(v, \tau)$ for all three problems is not a good idea. Why don't we
%  use $(v_i,\tau_i)$ introduced above? Things are much clearer this
%  way?}
\subsubsection{A mesh-dependent test norm}
\seclab{sec:meshDep}
Ideally, we would be interested in the use of the test norm 
\[
\|\left(v,\tau\right)\|_{V}^2 = \|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau\|^2 + \frac{1}{\epsilon}\|\tau\|^2
\]
for practical computations. However, the presence of the term $\|v\|$ together with $\sqrt{\epsilon}\|\grad v\|$ (and similarly $\|\div \tau\|$ and $\frac{1}{\sqrt{\epsilon}}\|\tau\|$ terms) induces boundary layers in the optimal test functions for under-resolved meshes. We can see this by recovering the strong form of the variational problem defining test functions. We first note that the variational problems for the $v$ and $\tau$ components of optimal test functions decouple from each other under this test norm. Then, examining the variational problem for the $v$ component only of an optimal test function, and assuming $\div \beta = 0$ for illustrative purposes, we have 
\begin{align*}
\left(\left(v,0\right),\left(\delta v,\delta\tau\right)\right)_{V} &= (v,\delta v) + \epsilon \left(\grad v,\grad \delta v\right) + \left(\beta \cdot \grad v, \beta \cdot \grad \delta v\right)\\
&= \left(v - \epsilon \Delta v - \div \left(\left(\beta \otimes \beta\right) \grad v\right), \delta v\right)_{L^2} + \langle \epsilon \grad v\cdot n, \delta v\rangle + \langle n\cdot \left(\beta\otimes\beta\right) \grad v , \delta v\rangle.
\end{align*}
After integration by parts, we recover the strong form of the operator $L$ inducing such a variational problem 
\[
Lv \coloneqq v - \epsilon \Delta v - \div \left(\left(\beta \otimes \beta\right) \grad v\right),
\]
where we neglect the resulting boundary terms from integration by parts for now.

The streamline direction $\beta$ induces an anisotropic diffusion, while the $\sqrt{\epsilon}\|\grad v\|_{L^2}$ term induces a small isotropic diffusion contribution everywhere. Since any vector in the cross-stream direction is in the null space of the anisotropic diffusion tensor, in the cross-stream directions, the optimal test function is governed only by the cross-stream part of the operator $L$
\[
L_{\rm \beta^\perp} \coloneqq v - \epsilon \Delta v,
\]
and can develop boundary layers in those directions. The presence of boundary layers has been verified through numerical computation as well; using an $H^1$-conforming finite element code with $hp$-adaptivity \cite{demkowicz2006computing}, the solution to the variational problem defining the optimal test function under the above test norm was computed. Figure~\ref{fig:boundaryTest} shows the result of such a computation for the $v$ component of an optimal test function under the above test norm. 
\begin{figure}[!h]
\centering
\subfigure{\includegraphics[scale=.2]{figs/mesh1.png}}
\subfigure{\includegraphics[scale=.2]{figs/sol1.png}}
\caption{The $v$ component of the optimal test function corresponding to flux $\widehat{u} = x(1-x)$ on the bottom side of a unit element for $\epsilon = 0.01$. The corresponding $hp$-mesh used to compute the solution is displayed to the left.}
\label{fig:boundaryTest}
\end{figure}
To avoid boundary layers in the optimal test functions, we follow \cite{DPGrobustness} in scaling the $L^2$ contributions of $v$ by $C_v(K)$, such that, when transformed to the reference element, both $C_v(K)\|v\|^2$ and $\epsilon\|\grad v\|^2$ are of the same magnitude. Similarly, we scale the $L^2$ contributions of $\tau$ by $C_\tau(K)$ such that $\frac{C_\tau(K)}{\epsilon} \|\tau\|^2$ and $\|\div \tau\|^2$ are of the same magnitude as well. For this work, we consider only isotropic refinements on quadrilateral elements in 2D.

@@

\subsubsection{Equivalence of energy norm with $\nor{\cdot}_U$}
%\subsection{Relations between trial and test constants}
\label{sec:main_bounds}

The main theoretical result of this work can now be given:
\begin{lemma}
Under the mesh-dependent test norm
\[
\|\left(v,\tau\right)\|_{V,\Oh}^2 = \|C_vv\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau\|^2 + \|C_\tau\tau\|^2,
\]
where $C_v, C_{\tau}\in L^2(\Omega)$ are defined elementwise through
\begin{align*}
\left.C_v\right |_K &= \min\left\{\sqrt{\frac{\epsilon}{|K|}},1\right\}\\
\left.C_{\tau}\right |_K &= \min\left\{\frac{1}{\sqrt{\epsilon}},\frac{1}{\sqrt{|K|}}\right\}.
\end{align*}
If $\beta$ satisfies \eqref{a_req}, \eqref{b_req}, and \eqref{c_req}, the DPG energy norm $\nor{\cdot}_E$ satisfies the following equivalence relations
\begin{align*}
\nor{u}_{L^2} + \nor{\sigma}_{L^2} + \epsilon\nor{\widehat{u}} + \sqrt{\epsilon}\nor{\widehat{f}_n} 
&\lesssim \nor{\left(u,\sigma, \widehat{u},\widehat{f}_n\right)}_E \\ 
%\nor{\left(u,\sigma, \widehat{u},\widehat{f}_n\right)}_E &\lesssim \nor{u}_{L^2} + \nor{C_\sigma \sigma}_{L^2} + \frac{1}{\sqrt{\epsilon}} \left(\nor{\widehat{u}} + \nor{\widehat{f}_n}\right),
\nor{\left(u,\sigma, \widehat{u},\widehat{f}_n\right)}_E &\lesssim \nor{u}_{L^2} + \nor{\frac{1}{\epsilon C_\tau} \sigma}_{L^2} + \frac{1}{\sqrt{\epsilon}} \left(\nor{\widehat{u}} + \nor{\widehat{f}_n}\right).
\end{align*}
%where $C_\sigma \in L^2(\Omega)$ is defined elementwise through $\left.C_\sigma\right|_K  = \max\left\{ \frac{1}{\sqrt{\epsilon}} , \frac{\sqrt{|K|}}{\epsilon} \right\}$.
\end{lemma}

@@

\subsubsection{Comparison of boundary conditions}

@@

The adjoint problem corresponding to Lemma~\ref{lemma_stream} in Section~\ref{sec:strategy3} is likewise reduced in 1D to the scalar equation
\begin{align}
\epsilon v_1'' + \beta v_1' &= -g \label{adjoint_1D}
\end{align}
with $v_1\in H^1_0\left((0,1)\right)$.  After multiplying this equation by $\beta v_1'$ and integrate by parts over $\Oh$, we can apply Young's inequality to get
%\[
%\int_0^1 \epsilon v_1'' \beta v_1' + \|\beta v_1'\|_{L^2}^2 = -\int_0^1 g \beta v_1'.\
%]
%Integrating by parts and Young's inequality gives
\begin{align*}
\left.\frac{\epsilon}{2} \beta v_1'^2\right|_0^1 + \|\beta v_1'\|_{L^2}^2 &\leq \frac{1}{2}\|g\|^2 + \frac{1}{2}\|\beta v_1'\|^2,
\end{align*}
implying that
\begin{align*}
\|\beta v_1'\|_{L^2}^2 & \lesssim \|g\|^2 + {\beta \epsilon} v_1'(0)^2.
\end{align*}
Let us restrict ourselves to the cases where $v_1$ is sufficiently smooth for $v'(0)$ to be well defined.  Taking $g=1$ (corresponding to a piecewise constant approximation) we can solve \eqref{adjoint_1D} exactly. 
\begin{figure}[!h]
\centering
\includegraphics[scale=.5]{figs/testBoundary1D.png}
\caption{$v_1(x) = \frac{e^{-\frac{x}{\epsilon}}}{e^{\frac{1}{\epsilon}}-1}\left(e^{\frac{1}{\epsilon}}\left(e^{\frac{x}{\epsilon}}-1\right) + \left(e^{\frac{1}{\epsilon}}-1\right)e^{\frac{x}{\epsilon}}x\right)$, the solution to the adjoint equation for $f=0$ and constant $\beta$ and load $g$ for $\epsilon = .01$.}
\label{fig:testBoundary1D}
\end{figure}
The solution $v_1$ is plotted in Figure~\ref{fig:testBoundary1D}, where we can see that $v_1(x)$ develops strong boundary layers of width $\epsilon$ near the inflow boundary $x=0$. Consequently, $\frac{\epsilon}{2} v_1'(0)^2 \approx \epsilon^{-1}$. Thus, we cannot conclude $\|\beta v'\| \lesssim \|g\|$ when $g$ is a constant,\footnote{Unlike the case of Dirichlet boundary conditions, the inflow condition on $ \widehat{f}_n = u(0)-\epsilon u'(0)$ induces an adjoint boundary condition $\tau(0)=0$, or equivalently $v'(0) = 0$, removing the non-robust term from the estimate.} and as a consequence cannot conclude that the robust error bound $\|u-u_h\|_{L^2} \lesssim \|(u,\sigma,\widehat{u},\widehat{f}_n)-(u_h,\sigma_h,\widehat{u}_h,\widehat{f}_{n,h})\|_E$ holds for the solution $u_h$.  More detailed 1D error bounds for Dirichlet boundary conditions are provided in \cite{DPG3}, and indicate the same lack of robustness under the test norm used thus far.\footnote{Demkowicz and Heuer proved in \cite{DPGrobustness} that for Dirichlet boundary conditions, robustness as $\epsilon \rightarrow 0$ is achieved by the test norm
\[
\|\left(\tau, v\right)\|_{V,w}^2 = \|v\| + \epsilon \|\grad v\| + \|\beta \cdot \grad v\|_{w+\epsilon} + \| \div \tau\|_{w+\epsilon} + \frac{1}{\epsilon}\|\tau\|_{w+\epsilon}
\]
where $\|\cdot \|_{w+\epsilon}$ is a weighted $L^2$ norm, where the weight $w \in (0,1)$ is required to vanish on $\Gamma_-$ and satisfy $\grad w = O(1)$. The need for this weight is necessary to account for the loss of robustness at the inflow.}

@@

\begin{figure}[h!]
\centering
\subfigure{
\includegraphics[scale=.36]{figs/roundoff_rates.png}
}
\subfigure{
\includegraphics[scale=.36]{figs/roundoff_ratio.png}
}
\caption{Energy error and $L^2$/energy error ratio for $\epsilon=1e-5$, $\epsilon=1e-6$, $\epsilon=1e-7$. Non-monotonic behavior of the energy error indicates conditioning issues and roundoff effects.}
\label{roundoff_figures}
\end{figure}

@@

\begin{figure}[h!]
\centering
\subfigure{
\includegraphics[scale=.36]{figs/rates_disc.png}
}
\subfigure{
\includegraphics[scale=.36]{figs/ratio_disc.png}
}
\caption{$L^2$ and energy errors, and their ratio for $\epsilon=.01$, $\epsilon=.001$, $\epsilon=.0001$, with discontinuous $u_0$ approximated by a Fourier expansion. }
\label{disc_sol_fourier}
\end{figure}
